---
title: 'SQL Exercise #1: Querying Movie Data - Solutions without Code'
output:
  html_document:
    df_print: paged
---

This notebook contains the **solutions without code** for Exercise 1 - Querying Movie Data.

```{r setup}
library(DBI)
db = dbConnect(RSQLite::SQLite(), dbname = "movies/data_movies.db")
knitr::opts_chunk$set(connection = "db")
```

## Question 1

Find the titles of all movies directed by Steven Spielberg.

```{sql, echo = FALSE}
SELECT title
FROM Movie
WHERE director = "Steven Spielberg";
```


## Question 2

Find all years that have a movie that received a rating of 4 or 5, and sort them in increasing order.

```{sql, echo = FALSE}
SELECT distinct Movie.year 
FROM Rating LEFT JOIN Movie
ON Rating.mID = Movie.mID
WHERE Rating.stars > 3
ORDER BY Movie.year;
```

## Question 3

Find the titles of all movies that have no ratings.

```{sql, echo = FALSE}
SELECT Movie.title
FROM Movie LEFT JOIN Rating
	ON Movie.mID = Rating.mID
WHERE Rating.stars IS NULL;
```

## Question 4

Some reviewers didn't provide a date with their rating. Find the names of all reviewers who have ratings with a NULL value for the date.

```{sql, echo = FALSE}
SELECT DISTINCT Reviewer.name 
FROM Rating LEFT JOIN Reviewer
	ON Rating.rID = Reviewer.rID
WHERE Rating.ratingDate IS NULL;
```

## Question 5

Write a query to return the ratings data in a more readable format: reviewer name, movie title, stars, and ratingDate. Also, sort the data, first by reviewer name, then by movie title, and lastly by number of stars.

```{sql, max.print = 20, echo = FALSE}
SELECT Reviewer.name, Movie.title, Rating.stars, Rating.ratingDate
FROM (Rating LEFT JOIN Reviewer ON Rating.rID = Reviewer.rID) LEFT JOIN Movie ON Rating.mID = Movie.mID
ORDER BY name ASC, title ASC, stars DESC;
```

## Question 6

For all cases where the same reviewer rated the same movie twice and gave it a higher rating the second time, return the reviewer's name and the title of the movie.

```{sql, echo = FALSE}
SELECT Reviewer.name, Movie.title
FROM (Rating AS r1 LEFT JOIN Movie ON r1.mID = Movie.mid) LEFT JOIN Reviewer ON r1.rID = Reviewer.rID, Rating AS r2
WHERE r1.rID = r2.rID AND 
	r1.mID = r2.mID AND
	r1.ratingDate > r2.ratingDate AND
	r1.stars > r2.stars;
```

## Question 7

For each movie that has at least one rating, find the highest number of stars that movie received. Return the movie title and number of stars. Sort by movie title.

```{sql, echo = FALSE}
SELECT title, MAX(stars) AS stars
FROM Rating LEFT JOIN Movie ON Rating.mID = Movie.mID
GROUP BY Rating.mID
ORDER BY title;
```

## Question 8

For each movie, return the title and the 'rating spread', that is, the difference between highest and lowest ratings given to that movie. Sort by rating spread from highest to lowest, then by movie title.

```{sql, echo = FALSE}
SELECT title, MAX(stars) - MIN(stars) AS spread
FROM Rating LEFT JOIN Movie on Rating.mID = Movie.mID
GROUP BY Rating.mID
ORDER BY spread DESC, title ASC;
```

## Question 9

Find the difference between the average rating of movies released before 1980 and the average rating of movies released after 1980. (Make sure to calculate the average rating for  each movie, then the average of those averages for movies before 1980 and movies after. Don't just calculate the overall average rating before and after 1980.)

```{sql, echo = FALSE}
SELECT AVG(before_1980.avg_per_movie) - AVG(after_1980.avg_per_movie)
FROM (SELECT AVG(stars) AS avg_per_movie
	FROM Rating LEFT JOIN Movie on Rating.mID = Movie.mID
	WHERE year < 1980
	GROUP BY movie.mID) AS before_1980,
	(SELECT AVG(stars) AS avg_per_movie
	FROM Rating LEFT JOIN Movie on Rating.mID = Movie.mID
	WHERE year > 1980
	GROUP BY movie.mID) AS after_1980;
```

```{r, echo = FALSE}
dbDisconnect(db)
```