---
title: 'SQL Exercise #4: Querying Social Data, Extra Exercises - Solutions without Code'
output:
  html_document:
    df_print: paged
---

This notebook contains the **solutions without code** for Exercise 4 - Querying Social Data, Extra Exercises.

```{r setup}
library(DBI)
db = dbConnect(RSQLite::SQLite(), dbname = "social/data_social.db")
knitr::opts_chunk$set(connection = "db")
```

## Question 1

For every situation where student A likes student B, but student B likes a different student C, return the names and grades of A, B, and C.

```{sql, echo = FALSE}
SELECT DISTINCT hs1.name, hs1.grade, hs2.name, hs2.grade, hs3.name, hs3.grade 
FROM Highschooler AS hs1, Highschooler AS hs2, Highschooler AS hs3, Likes AS l1, Likes AS l2
WHERE l1.ID1 = l2.ID2 AND
    l1.ID2 != l2.ID1 AND
	l1.ID1 = hs2.ID AND
	l1.ID2 = hs3.ID AND
	l2.ID1 = hs1.ID
```

## Question 2

Find those students for whom all of their friends are in different grades from themselves. Return the students' names and grades.

```{sql, echo = FALSE}
SELECT DISTINCT hs1.name, hs1.grade
FROM Friend, Highschooler AS hs1
WHERE hs1.grade NOT IN (
    SELECT hs2.grade
    FROM Friend, Highschooler AS hs2
    WHERE Friend.ID1 = hs1.ID AND
        Friend.ID2 = hs2.ID
)
```

## Question 3

What is the average number of friends per student? (Your result should be just one number.)

```{sql, echo = FALSE}
SELECT AVG(friend_count)
FROM (SELECT COUNT(ID2) AS friend_count
    FROM Friend
    GROUP BY ID1)
```

## Question 4

Find the number of students who are either friends with Cassandra or are friends of friends of Cassandra. Do not count Cassandra, even though technically she is a friend of a friend.

```{sql, echo = FALSE}
SELECT COUNT(*)
FROM (
    SELECT Highschooler.ID
    FROM Friend, Highschooler
    WHERE Friend.ID1 = (SELECT ID FROM Highschooler WHERE name = 'Cassandra') AND
        Friend.ID2 = Highschooler.ID
    UNION ALL
    SELECT Highschooler.ID
    FROM Friend, Highschooler
    WHERE Friend.ID1 IN (
        SELECT Highschooler.ID
        FROM Friend, Highschooler
        WHERE Friend.ID1 = (SELECT ID FROM Highschooler WHERE name = 'Cassandra') AND
            Friend.ID2 = Highschooler.ID
        ) AND
        Friend.ID2 = Highschooler.ID AND
        Highschooler.ID != (SELECT ID FROM Highschooler WHERE name = 'Cassandra')
    )
```

## Question 5

Find the name and grade of the student(s) with the greatest number of friends.

```{sql, echo = FALSE}
SELECT name, grade
FROM Highschooler
WHERE ID IN (
    SELECT ID1
        FROM Friend
        GROUP BY ID1
        HAVING COUNT(ID2) = (
            SELECT MAX(friend_count) AS friends_max
            FROM (SELECT COUNT(ID2) AS friend_count
                FROM Friend
                GROUP BY ID1)
        )
)
```

```{r, echo = FALSE}
dbDisconnect(db)
```