---
title: 'SQL Exercise #5: Modifying Movie Data - Solutions without Code'
output:
  html_document:
    df_print: paged
---

This notebook contains the **solutions without code** for Exercise 5 - Modifying Movie Data.

```{r setup}
library(DBI)
library(RSQLite)

# Mechanism to prevent permanent edits of the database:
# 1) Creating an in-memory temporary database and loading the reference database
# 2) Copying the static database to the temporary in-memory database
# 3) Closing the connection to the static database
# 4) Performing all operations on the in-memory database only

db_temp <- dbConnect(RSQLite::SQLite(), ":memory:")
db_static <- dbConnect(RSQLite::SQLite(), dbname = "movies/data_movies.db")

RSQLite::sqliteCopyDatabase(from = db_static, to = db_temp)
dbDisconnect(db_static)

knitr::opts_chunk$set(connection = "db_temp")
```

## Question 1

Add the reviewer Roger Ebert to your database, with an rID of 209.

```{sql, echo = FALSE}
INSERT INTO Reviewer
VALUES(209, 'Roger Ebert');
```

```{sql, echo = FALSE}
-- Verify that it works.

SELECT * 
FROM Reviewer;
```


## Question 2

Insert 5-star ratings by James Cameron for all movies in the database. Leave the review date as NULL.

```{sql, echo = FALSE}
INSERT INTO Rating
(rID, mID, stars, ratingDate)
    SELECT Reviewer.rID, Movie.mID, 5 AS stars, NULL AS ratingDate
    FROM Movie, Reviewer
    WHERE Reviewer.name = 'James Cameron';
```

```{sql, echo = FALSE}
-- Verify that it works.

SELECT * 
FROM Rating
WHERE rID IN (SELECT rID From Reviewer WHERE name = 'James Cameron');

```

## Question 3

For all movies that have an average rating of 4 stars or higher, add 25 to the release year. (Update the existing tuples; don't insert new tuples.)

```{sql, echo = FALSE}
UPDATE Movie
SET year = year + 25
WHERE mID IN (
    SELECT mid
    FROM Rating
    GROUP BY mID
    HAVING AVG(stars) >= 4.0);
```

```{sql, echo = FALSE}
-- Verify that it works.

SELECT Rating.mID, Movie.year, AVG(Rating.stars)
FROM Rating, Movie
WHERE Rating.mID = Movie.mID
GROUP BY Rating.mID
HAVING AVG(Rating.stars) >= 4.0;
```

## Question 4

Remove all ratings where the movie's year is before 1970 or after 2000, and the rating is fewer than 4 stars.

```{sql, echo = FALSE}
DELETE FROM Rating
WHERE stars < 4 AND
    mid IN (SELECT mid
            FROM Movie
            WHERE (year > 2000 OR year < 1970));
```

```{sql, echo = FALSE}
-- Verify that it works.

SELECT *
FROM Rating
WHERE stars < 4 AND
    mid IN (SELECT mid
        FROM Movie
        WHERE (year > 2000 OR year < 1970));
```

```{r, echo = FALSE}
dbDisconnect(db_temp)
```
