---
title: 'SQL Exercise #6: Modifying Social Data - Solutions with Code'
output:
  html_document:
    df_print: paged
---

This notebook contains the **solutions with code** for Exercise 6 - Modifying Social Data.

```{r setup}
library(DBI)
library(RSQLite)

# Mechanism to prevent permanent edits of the database:
# 1) Creating an in-memory temporary database and loading the reference database
# 2) Copying the static database to the temporary in-memory database
# 3) Closing the connection to the static database
# 4) Performing all operations on the in-memory database only

db_temp <- dbConnect(RSQLite::SQLite(), ":memory:")
db_static <- dbConnect(RSQLite::SQLite(), dbname = "social/data_social.db")

RSQLite::sqliteCopyDatabase(from = db_static, to = db_temp)
dbDisconnect(db_static)

knitr::opts_chunk$set(connection = "db_temp")
```

## Question 1

It's time for the seniors to graduate. Remove all 12th graders from Highschooler.

```{sql}
DELETE FROM Highschooler
WHERE grade = 12;
```

```{sql}
-- Verify that it works.

SELECT MAX(grade)
FROM Highschooler;
```


## Question 2

If two students A and B are friends, and A likes B but not vice-versa, remove the Likes tuple.

```{sql}
DELETE FROM Likes
WHERE Likes.ID1 IN (
        SELECT Friend.ID2
        FROM Friend
        WHERE Friend.ID1 = Likes.ID2) AND
    Likes.ID2 IN (
        SELECT l.ID2
        FROM Likes AS l
        WHERE Likes.ID1 = l.ID1) AND
    Likes.ID1 NOT IN (
        SELECT l.ID2
        FROM Likes AS l
        WHERE Likes.ID2 = l.id1);
```

```{sql}
-- Verify that it works.

SELECT *
FROM Likes
WHERE Likes.ID1 IN (
        SELECT Friend.ID2
        FROM Friend
        WHERE Friend.ID1 = Likes.ID2) AND
    Likes.ID2 IN (
        SELECT l.ID2
        FROM Likes AS l
        WHERE Likes.ID1 = l.ID1) AND
    Likes.ID1 NOT IN (
        SELECT l.ID2
        FROM Likes AS l
        WHERE Likes.ID2 = l.id1);
```

```{sql}
-- Verification via the reference solution that was used in the course.

SELECT H1.name, H1.grade, H2.name, H2.grade 
FROM Likes L, Highschooler H1, Highschooler H2
WHERE L.ID1 = H1.ID and L.ID2 = H2.ID order by H1.name, H1.grade;
```


## Question 3

For all cases where A is friends with B, and B is friends with C, add a new friendship for the pair A and C. Do not add duplicate friendships, friendships that already exist, or friendships with oneself. (This one is a bit challenging; congratulations if you get it right.)

```{sql}
INSERT INTO Friend (id1, id2)
    SELECT DISTINCT f1.ID1 AS id1, f3.ID1 AS id2
    FROM Friend AS f1, Friend AS f2, Friend AS f3
    WHERE f2.ID1 IN (
        SELECT Friend.ID2
        FROM Friend
        WHERE f1.ID1 = Friend.ID1
    ) AND
        f2.ID1 IN (
        SELECT Friend.ID2
        FROM Friend
        WHERE f3.ID1 = Friend.ID1
        ) AND 
        f3.ID1 NOT IN (
        SELECT Friend.ID2
        FROM Friend
        WHERE f1.ID1 = Friend.ID1
        )
        AND (f1.ID1 != f3.ID1);
```

```{sql}
-- Verification via the reference solution that was used in the course.

SELECT ID, name, grade,
    (SELECT count(*)
        FROM Friend
        WHERE id1 = H.id)
FROM Highschooler H
ORDER BY ID;

```

```{r, echo = FALSE}
dbDisconnect(db_temp)
```
